---
title: 'Tutorial: functions for biocultural complexity'
author: "M. Humberto Reyes-Valdés"
date: "2/6/2019"
output: pdf_document
classoption: landscape
---

This tutorial shows the use of the functions coded in the script *biocultural.R*, to analyzie biocultural complexity from a table of cultures and species usage. In this case, I use the file *toyset.csv*, which is an artificial simple data set. The platform and language is R.

1. Open an R console

2. Read the data set


```{r}
dat<-read.csv("data/toyset.csv",head=T)
```

3. Visualize data features

```{r}
dat
#rep is the representation frequency of cultural groups. 
#In this case the assumption is that all groups are equally represented,
#and one can write 1 for each entry. 
#culture is the column of cultural group names
#sp1 to sp5 are species names, with columns
#representing the observed field counts
str(dat)
```

4. Source script

```{r}
source("biocultural.R")
```

5. The function bc.info creates an object with basic estimated biocultural parameters

```{r}
ob.sim<-bc.info(dat)
names(ob.sim)
```

6. A more complete and practical function is *bc.tables*, which performs bootstrap for bias correction, and estimates standard errors and confidence intervals. Definitions for table columns are in the respective paper (in review).

```{r}
#Run bc.tables with 1000 bootstrap resamplings
set.seed(123)
x<-bc.tables(dat,1000)
names(x)
```

7. Display table for species

```{r}
library(knitr)
kable(x$tabSpecies)
```

8. Display table for cultural groups
```{r}
kable(x$tabCultures)
```

9. Display a table for Biocultural Complexity and Relative Biocultural Complexity , and another one with confidence intervals (see See Bryan FJ Manly, Randomization, Bootstrap and Monte Carlo methods in Biology, 2006, page 44).

```{r}
#Biocultural Complexity and its relative counterpart
kable(x$tabBC)
#Confidence intervals
kable(x$taBC_CI_2)
#Note: the attribute taBC_CI_1 contains a table with confidence intervals, 
#based solely on quantiles. 
#taBC_CI_2 gives intervals by the second method described by Manly (2006), 
#which appears to work better for this application.
```


10. Perform a chi square test for association between cultures and species. A montecarlo simulation is used to get the *p* value.

```{r}
chisq.test(dat[-c(1,2)],simulate.p.value=T)
```
